Appendix S4. The relative importance of evolution in populations and communities

### 2024-11-10 manuscript/Appendix_S4_v01.Rmd

m.11.hmsc = Hmsc(Y = Y, XData = XData, XFormula = ~E+Esq+dx1, studyDesign = studyDesign, ranLevels = list(sample = rL))
# Bayesian model parameters
nChains <- 2
thin <- 5
samples <- 2000
transient <- 1000*thin
verbose <- 500*thin
# sample MCMC
m.11.sample <- sampleMcmc(m.11.hmsc,thin=thin,sample=samples,transient=transient,nChains=nChains,verbose=verbose)
m11.post.hmsc <- convertToCodaObject(m.11.sample)
```

```{r warning=FALSE}
VP11 <- computeVariancePartitioning(m.11.sample,group=c(1,1,1,2),groupnames=c("Env","x1"))
plotVariancePartitioning(m.11.sample,VP11,cols=c("white","skyblue","darkgrey"),args.legend=list(cex=0.75,bg="transparent"))
```
```{r}
# prepare for plot
pred2 <- predict(m.11.sample)
hold1 <- list()
hold2 <- list()
for(i in 1:length(N0)){
  hold1[[i]] <- as.numeric(Y[,i])
  data <- lapply(pred2, function(x) x[, colnames(x)[i], drop = FALSE])
  hold2[[i]] <- matrix(unlist(data), nrow=length(pred2), byrow=TRUE)
}
y <- unlist(hold1) # vector of each species' observed N across time points
yrep <- as.matrix(data.frame(hold2))
### Mod2 plot ###
par(mgp=c(2,0.45,0), tcl=-0.4, mar=c(1.3,1.2,0,0))
color_scheme_set("brightblue")
#ppc_dens_overlay(y, yrep[1:50, ])
ppc_intervals(y,yrep,x=dat$time[1:(t-1)],prob = 0.95) +
  labs(x = "time",y = "ln(N)",) +
  theme(plot.margin = unit(c(0, 0, 0, 0), "cm")) +
  theme(legend.position = "none")
```

@@

>>>>

m.11.hmsc = Hmsc(Y = Y, XData = XData, XFormula = ~E+Esq+dx1, studyDesign = studyDesign, ranLevels = list(sample = rL))
# Bayesian model parameters
nChains <- 2
thin <- 5
samples <- 2000
transient <- 1000*thin
verbose <- 500*thin
# sample MCMC
m.11.sample <- sampleMcmc(m.11.hmsc,thin=thin,sample=samples,transient=transient,nChains=nChains,verbose=verbose)
m11.post.hmsc <- convertToCodaObject(m.11.sample)
```
```{r}
m11.post.hmsc <- convertToCodaObject(m.11.sample)
summary(m11.post.hmsc$Beta)
bayesplot::mcmc_trace(m11.post.hmsc$Beta)
bayesplot::mcmc_areas(m11.post.hmsc$Beta,area_method = c("equal height"))
```
```{r warning=FALSE}
VP11 <- computeVariancePartitioning(m.11.sample,group=c(1,1,1,2),groupnames=c("Env","x1"))
plotVariancePartitioning(m.11.sample,VP11,cols=c("white","skyblue","darkgrey"),args.legend=list(cex=0.75,bg="transparent"))
```
```{r}
# prepare for plot
pred2 <- predict(m.11.sample)
hold1 <- list()
hold2 <- list()
for(i in 1:length(N0)){
  hold1[[i]] <- as.numeric(Y[,i])
  data <- lapply(pred2, function(x) x[, colnames(x)[i], drop = FALSE])
  hold2[[i]] <- matrix(unlist(data), nrow=length(pred2), byrow=TRUE)
}
y <- unlist(hold1) # vector of each species' observed N across time points
yrep <- as.matrix(data.frame(hold2))
### Mod2 plot ###
par(mgp=c(2,0.45,0), tcl=-0.4, mar=c(1.3,1.2,0,0))
color_scheme_set("brightblue")
#ppc_dens_overlay(y, yrep[1:50, ])
ppc_intervals(y,yrep,x=dat$time[1:(t-1)],prob = 0.95) +
  labs(x = "time",y = "ln(N)",) +
  theme(plot.margin = unit(c(0, 0, 0, 0), "cm")) +
  theme(legend.position = "none")
```

@@

We can see that the variation in Y is a combination of the relative importance of density-dependence and of trait evolution. We can see that relative importance changes across $h^2$ levels. What drives changes in this relative importance?

```{r}
d_dat <- gdat[gdat$time != 1,c(1,2,3,5)]
d_dat <- as.data.frame(d_dat)
d_dat$dxt <- NA
d_dat$ddt <- NA
d_dat$dN <- NA
d_dat$dd <- NA
for(i in 1:11){
  sub <- gdat[gdat$h2==levels(gdat$h2)[i],]
  d_dat$dxt[d_dat$h2==levels(d_dat$h2)[i]] <- abs(sub$x1[2:60]-sub$x1[1:59]) # absolute value of change in trait value
  d_dat$ddt[d_dat$h2==levels(d_dat$h2)[i]] <- alpha[1,1]*sub$N1[2:60] - alpha[1,1]*sub$N1[1:59]
  d_dat$dN[d_dat$h2==levels(d_dat$h2)[i]] <- sub$N1[2:60] - sub$N1[1:59]
  d_dat$dd[d_dat$h2==levels(d_dat$h2)[i]] <- alpha[1,1]*d_dat$N1[d_dat$h2==levels(d_dat$h2)[i]]
}
```
```{r}
### Plot of time xx ###
d_dat <- d_dat[d_dat$h2 != 0,]
d_dat$ln_N <- log(d_dat$N1)
ggplot2::ggplot(d_dat, aes(dxt, dN, col=h2)) + geom_line() + geom_point()
ggplot2::ggplot(d_dat, aes(ddt, dN, col=h2)) + geom_line() + geom_point()
### ###
```
